---
title: "ST558 - Project 2"
author: 'Group 9: Max Marion-Spencer & Xin Wang'
date: '2022-07-03'
output: html_document
params: 
      data_channel: bus
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
library(tibble)
library(knitr)
library(dplyr)
library(ggplot2)
library(corrplot)
```

## Introduction 

note: 
max, this is the variables I used, feel free to add more if needed

2. n_tokens_title: Number of words in the title 
3. n_tokens_content: Number of words in the content 
7. num_hrefs: Number of links 
9. num_imgs: Number of images 
10. num_videos: Number of videos 
12.num_keywords: Number of keywords in the metadata 

13. data_channel_is_lifestyle: Is data channel 'Lifestyle'? 
14. data_channel_is_entertainment: Is data channel 'Entertainment'? 
15. data_channel_is_bus: Is data channel 'Business'? 
16. data_channel_is_socmed: Is data channel 'Social Media'? 
17. data_channel_is_tech: Is data channel 'Tech'? 
18. data_channel_is_world: Is data channel 'World'? 

31. weekday_is_monday: Was the article published on a Monday? 
32. weekday_is_tuesday: Was the article published on a Tuesday? 
33. weekday_is_wednesday: Was the article published on a Wednesday? 
34. weekday_is_thursday: Was the article published on a Thursday? 
35. weekday_is_friday: Was the article published on a Friday? 
36. weekday_is_saturday: Was the article published on a Saturday? 
37. weekday_is_sunday: Was the article published on a Sunday? 
38. is_weekend: Was the article published on the weekend? 


## Data

```{r}
# read in data
news <- read_csv("OnlineNewsPopularity.csv") %>% 
  as_tibble() %>% select(c(3,4,8,10,11,13,14:19,32:39,61))
colnames(news)
news
```

```{r}
## pivot data channel from wide to long 
news <- news %>%
  pivot_longer(
    cols = starts_with("data_channel"),
    names_to = "data_channel",
    values_to = "data_channel_type"
  ) %>%
  pivot_longer(
    cols = starts_with("weekday_is"),
    names_to = "weekday",
    values_to = "weekday_type"
  )

##remove prefix and filter both type with 1
news<- news %>% 
  mutate(weekday = str_sub(weekday, 12, -1),
          data_channel = str_sub(data_channel, 17, -1)) %>% 
  filter(data_channel_type == 1 & weekday_type ==1)
```

```{r}
## factor weekday and data channel
news$weekday <- as.factor(news$weekday)
news$weekday<- ordered(news$weekday, levels=c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday"))
levels(news$weekday)

news$data_channel <- as.factor(news$data_channel)
levels(news$data_channel)
```

```{r}
# filter out one data channel 
news <- news %>% filter(data_channel == params$data_channel)
news
```

## Summarizations




### Contingency Tables

This table shows the count of days in a week for article publish. It helps us to see which day of the week article published the most and whether there are more article published on the weekdays or weekends. 

```{r}
kable(t(table(news$weekday)), caption = "Article Published Day")
```

We transform shares into a binary variable using the decision threshold of 1,400. This contingency table shows the count of shares on the weekdays and weekends. 

```{r}
newsa <- news %>% 
  mutate(SharesClass = if_else(shares >=1400, "More than 1,400","Less than 1,400"), 
         is_weekend = if_else(is_weekend== 0, "Weekday", "Weekend"))

kable(table(newsa$SharesClass,newsa$is_weekend), caption = "Count of Shares")
```

### Correlation Plot

```{r}
cor.variables <- news %>% select(n_tokens_title, n_tokens_content, num_hrefs, num_imgs, num_videos,num_keywords)
correlation <- cor(cor.variables, method = "spearman")
corrplot(correlation)
```

## Modeling

## Comparison

## Automation

this should work, can test after finished above 

```{r,eval=FALSE}
data_channel <- c("bus","entertainment","lifestyle","socmed","tech","world")
output_file <- paste0(data_channel, "Analysis.md")
params = lapply(data_channel, 
                FUN = function(x){
                  list(data_channel = x)})
reports <- tibble(output_file, params)
apply(reports, 
      MARGIN = 1,
      FUN = function(x){
        rmarkdown::render(input = "st558 - project2.Rmd", 
                          output_format = "github_document", 
                          output_options = list(html_preview = FALSE), 
                          output_file = x[[1]], 
                          params = x[[2]])})
```

